---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---
[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/mqtt.svg?branch=master)](https://travis-ci.org/hrbrmstr/mqtt)

# mqtt

Interoperate with 'MQTT' Message Brokers

## Description

'MQTT' is a machine-to-machine ('M2M')/"Internet of Things" connectivity protocol. It was designed as an extremely lightweight publish/subscribe messaging transport. It is useful for connections with remote locations where a small code footprint is required and/or network bandwidth is at a premium. For example, it has been used in sensors communicating to a broker via satellite link, over occasional dial-up connections with healthcare providers, and in a range of home automation and small device scenarios. It is also ideal for mobile applications because of its small size, low power usage, minimised data packets, and efficient distribution of information to one or many receivers. Tools are provided to interoperate with 'MQTT' message brokers in R.

## What You Need To Get This Working

You need to install the [`mosquitto` libraries](https://mosquitto.org/download/) and ensure they're on your system `PATH` so R can find them. This will be easier in the future, but coarse for the moment.

For macOS, that's as easy as:

    brew install mosquitto
    
For Debian/Ubuntu, it will be something like:

    sudo apt install libmosquitto-dev libmosquittopp-dev

For Windows folks: install Linux or get a mac ;-) Seriously, I'm hoping to have support for that soon.

## NOTICE

I had neglected to commit the code that makes the default client-id unique. So if you had tried this just after the blog post and were getting wonky results, that's the cause. The default client id will always be unique, now, but you should use your own, unique client ids in production.

## Current Functionality

You can subscribe to a topic on a server over plaintext. No authentication methods are supported (yet) and no ability to use encryption exists (yet).

When you subscribe, you should pass in a "callback handler". Said handler should have the same "signature" as the built-in default one, which is:

    mqtt_default_message_callback <- function(id, topic, payload, qos, retain)
    
Those parameters are:

- `id`: the message id
- `topic`: the message topic
- `payload`: the message payload (raw)
- `qos`: the effective qos for the message
- `retain`: is this message marked as "retain"?

`payload` is a raw vector. The example below shows how to work with this data.

If you return "`quit`" from this function, the subscription will be closed and control return to R.

I can see that publishing data to an MQTT broker would be useful for R so that is on the TODO. 

PLEASE file an issue and/or PR if you have ideas in mind for this. I have more "evil" notions in mind (quantifiying expousure of sensitive data on public MQTT channels). I suspect others have "real" needs for this.

## What's Inside The Tin

The following functions are implemented:

- `mqtt_default_connection_callback`:	mqtt default connection callback function
- `mqtt_default_disconnection_callback`:	mqtt default disconnection callback function (does - `nothing`)
- `mqtt_default_message_callback`:	mqtt default message callback function
- `mqtt_silent_connection_callback`:	mqtt silent connection callback function (does nothing)
- `topic_subscribe`:	Subscribe to an MQTT Topic

## Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/mqtt")
```

```{r message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
options(width=120)
```

## Usage

```{r message=FALSE, warning=FALSE, error=FALSE}
library(mqtt)

# current verison
packageVersion("mqtt")

# internal function to see which mosquitto library is being used
print(mqtt:::mqtt_version())
```

### Live subtitles

For whatever reason, someone is using the public `test.mosquitto.org` plaintext broker to push out live subtitles. It's strangely mezmerizing to watch it slowly scroll by. Let's see the next 50 (as of the time this Rmd ran):

```{r}
# We are going to cap it at 50 so we have to initialize a global we'll update
x <- 0

# Now, we need a callback function. This will get called everytime we get a message.
# the `topic` string will be passed in so you can compare that quickly.
# the `payload` is a raw vector since this can be pretty strange data (esp if you 
# subscribe to a wildcard).
# 
# You can use `rawToChar()` if you know it's going to be safe, but `readBin()` is
# a tad safter. Ideally, the package will have some sanitizing functions to make
# this easier and more robust.
my_msg_cb <- function(id, topic, payload, qos, retain) {
  
  if (topic == "bbc/subtitles/bbc_news24/raw") { # when we see BBC msgs, we'll cat them
    x <<- x + 1
    cat(readBin(payload, "character"), "\n", sep="")
  } else {
    print(topic)
  }

  return(if (x==50) "quit" else "continue") # "continue" can be "". anything but "quit"
}

# now, we'll subscribe to a wildcard topic at `test.mosquitto.org` on port 1883. 
# those are defaults in `topic_subscribe()` to make it easier to have some quick
# wun with the package.
topic_subscribe(topic="bbc/subtitles/bbc_news24/raw", message_callback=my_msg_cb)
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.